Gametic Selection, Meiotic Drive, Sex Ratio Bias, and Transitions Between sex-determination systems

### 2017-07-06 LaTeX/Turnover.tex
region plot figure added, working through mscott comments on discussion

%\textcolor{blue}{
In Figure \ref{fig:regionplot_ 2a} we show a region of parameter space under which both neo-W haplotypes invade ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) when there is overdominance in females and no haploid selection \citep[corresponding to Figure 2a in][]{Otto2014}. 
Wherever both haplotypes have positive growth rates (gray region of Figure \ref{fig:regionplot_ 2a}), invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where only one haplotype can spread (white region of Figure \ref{fig:regionplot_ 2a}), a neo-W can invade as long as equation \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is small enough and yet still indicates that more loosely linked sex-determining regions can spread.
%}

@@

\textcolor{blue}{As with the discussion, this section might need some editing for (1) unclearly describing looser linkage never evolving without haploid selection, which we now show is possible above and (2) going on about sex ratio selection too much - this should be changed to something along the lines of  ``male biased zygotic sex ratios ($\zeta-1/2$) are of the order of $\epsilon^2$. Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section.
''
}

@@

%Sex ratios
Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Our model provides a good opportunity to determine whether Fisherian sex-ratio selection provides a useful explanation for the evolution of new sex-determining loci in other contexts. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
We will also disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
If the ancestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
Thus, if the zygotic sex ratio is crucial to the evolution of new genetic sex-determining systems, invasion into ZW and XY systems will be distinct. 
However, under weak selection we find that invasion by a homogametic neo-sex-determining allele (XY to XY or ZW to ZW) or by a heterogametic neo-sex-determining allele (XY to ZW or ZW to XY) occur under the same conditions. 
That is, we can show that $\lambda_{Y',XY}=\lambda_{W',ZW}$ and $\lambda_{Y',ZW}=\lambda_{W',XY}$ (at least up to order $\epsilon^3$; for a numerical example, compare Figure \ref{fig:Combination_Turnover}A,B to Figure \ref{fig:Combination_Turnover}C,D).
As it turns out, under weak selection the strength of sex-ratio selection favouring, say, the invasion of a neo-W in an XY system is the same as the strength of meiotic drive favouring the invasion of a neo-Y in a ZW system. 
Even when these forces are not exactly the same (e.g., under tight sex-linkage; compare black and red curves near -25 and 25cM in Figure \ref{fig:Combination_Centimorgans}), it is important to remember that sex-ratio selection is only one of many potential selective forces acting to determine transitions between sex-determining systems.
It is even possible for the other selective forces to overwhelm sex-ratio selection and favour sex-determination transitions that create sex-ratio biases (Figure \ref{fig:Combination_Turnover}A,C).

Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
We will also disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
We might therefore predict a difference when we  XY to ZW transitions 
However, under weak selection we find that invasion by a homogametic neo-sex-determining allele (XY to XY or ZW to ZW) or by a heterogametic neo-sex-determining allele (XY to ZW or ZW to XY) occur under the same conditions. 
This is because, with weak selection, male biased zygotic sex ratios ($\zeta-1/2$) favouring the invasion of a neo-W are of the same order as meiotic drive favouring the invasion of a driving ne $\epsilon^2$.

Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section

@@

We next consider the case where the new sex-determining mutation, $m$, causes sex to be determined probabilistically or by heterogeneous environmental conditions (environmental sex determiner, ESD). 
We assume that individuals carrying the $m$ allele develop as females with probability $k$ (e.g., in a fraction $k$ of the environments they randomly experience). 
Assuming weak selection, the spread of these mutations is given by

@@

\textcolor{blue}{maybe re-order to put the results up front and focus on sex ratio less. Somewhere around these first two paragraphs we should include our tight linkage results. At the moment it doesn't make total sense because it assumes that tighter linkage is always favoured unless there is haploid selection.}

%HAPLOID SELECTION CAN ACT LIKE SEXUALLY-ANTAGONISTIC SELECTION
It has previously been demonstrated that new sex-determining systems can evolve if there is genetic variation maintained by sexually-antagonistic selection \citep{vanDoorn:2007eu,vanDoorn:2010hu}. 
In particular, transitions to new sex-determining systems can occur when new sex-determining regions are more closely linked to a sexually-antagonistic locus.
Our results show that genetic variation at loci that experience haploid selection can also generate selection in favour of new sex-determining systems. 
New sex-determining alleles are again favoured if they are linked with a locus under haploid selection and the ancestral sex-determination locus is not.
However, with haploid selection, heterogametic transitions (XY to ZW or ZW to XY) can also occur when the new sex-determining region is less closely linked to the locus under selection.

%HAPLOID SELECTION ALSO ALLOWS LOOSER LINKAGE TO EVOLVE
Neo-W (neo-Y) alleles invade when their fitness in females (males) is greater than the mean fitness of females (males) under the ancestral sex-determination system.
With sexually-antagonistic selection (between diploid sexes) only, linkage between a selected locus and the sex-determining region strengthens associations between male beneficial alleles and the male-determining allele (Y or Z) and between female beneficial alleles and the female-determining allele (X or W). 
Thus, the mean fitness of both males and females increases with closer linkage to the sex-determining region. 
Therefore, new sex-determining alleles only invade if they are more closely linked than the ancestral sex-determining region. 
However, if there is haploid selection on loci linked to an XY (ZW) sex-determining region, selection can maintain polymorphisms at which the mean fitness of females (males) is lower than it would be without sex-linkage. 
In these cases, unlinked neo-W (neo-Y) alleles can increase female (male) fitness, at a cost to the other sex, and invade despite lowering mean fitness (Figure \ref{fig:Combination_MeanFit}).

%TIGHT LINKAGE RESULTS

\textcolor{blue}{maybe drop paragraph?}
\textcolor{red}{I messed with the sex-ratio selection paragraphs to tone down our "it doesn't matter" speech from before. Have at any amendments you'd like to make.}
%SEX-RATIO SELECTION
Linkage between haploid selected loci and sex-determining regions cases biased zygotic sex ratios \citep{Hamilton:1967ts,Burt:2006,Field:2012fd,Field:2013cc}.
One might then expect Fisherian sex-ratio selection to drive the spread of new sex-determining systems that bring the sex ratio closer to 50:50. 
Fisherian sex-ratio selection follows from the fact that, for an autosomal locus, half of the genetic material is inherited from a male, and half from a female \citep{Fisher:1930wy,West:2009we}. 
Thus, if the population sex ratio is biased towards females, the average per-individual contribution of genetic material to the next generation from males is greater than the contribution from females (and vice versa for male-biased sex ratios). 
Therefore, a mutant that increases investment in males will spread via the higher per-individual contributions made by males. 
%%%That is, under Fisherian sex-ratio selection, the success of a mutant relative to the non-mutant depends, in equal parts, on the contributions made by males and females to the next generation. 
%An implicit assumption of Fisherian sex-ratio selection is that the mutant allele is autosomal and has the same inheritance pattern as the non-mutant allele. 
%The mutations we consider here, neo-sex-determining alleles, break this assumption. 
%For example, the success of neo-Y/neo-W mutations depends only on the number of alleles contributed by males/females (Table \ref{tab:haplotype_growth}). 
%In this respect, a neo-W is similar to a cytoplasmic element, which also does not experience selection to balance sex ratios \citep{Frank:1989vl,Werren:1998co,Chase:2007bf}.
%Even mutants that are equally likely to be found in males or females, such as an environmental sex determination mutation (equation \ref{eq:lambda_ESD}), are not strictly autosomal if they determine sex. 
%Thus, despite the fact that sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal modifiers \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va}, we do not find evidence of Fisherian sex-ratio selection acting during invasion by neo-sex-determination systems (e.g., see Figure \ref{fig:Combination_Turnover} and \citealt{Ubeda:2015fx}, in which a neo-Y invades despite biasing sex ratios).

\textcolor{blue}{Maybe we could go along the angle of ``under the weak recombination assumptions, there is no asymmetry between XY to ZW and ZW to XY, indicating that sex ratio selection does not dominate. Asymmetry can be larger when linkage is tight (e.g., Fig 2C) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
Thus, haploid selection can favour heterogametic transitions both via sex ratio selection and via the fitness of alleles that are associated with the neo-sex-determining allele; these selection pressures are often of equal magnitude.''}

Sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va} and sex-linked \citep{Ubeda:2015fx} modifiers.
We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
For instance, when an allele that drives in males is linked to an XY locus it will often become associated with the Y and therefore produce a male bias ($\zeta<1/2$).
This male bias increases the potential for a neo-W to invade (as we then have $(2\zeta)^{-1}>1$, Table \ref{tab:haplotype_growth}), which can equalize the sex-ratio \citep[for a related example see][]{Ubeda:2015fx}.
%However, the neo-W may not equalize the sex-ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
However, this sex-ratio selection can be overwhelmed when the driving allele has additional selective effects (e.g., when it is detrimental to male diploids but beneficial for female diploids; Table \ref{tab:specialcases}), preventing the neo-W from invading. %and equalizing the sex ratio.
Conversely, these additional selective effects can even favour transitions between sex-determining systems that create new sex-ratio biases.
For example, in an ancestral ZW system, an allele that drives only in males can allow a linked neo-Y to invade, despite the fact it creates a male bias.
This of course generates new sex-ratio selection that may drive further turnover \citep{Ubeda:2015fx}.
What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.

@@

%HOW TO TEST EMPIRICALLY
\textcolor{blue}{I think Sally was sceptical of this idea. i.e., if the second term in \ref{eq:lambda_neoW} is negative, then get less transitions with haploid selection included, right? }
Taken at face value, our results indicate that transitions in heterogametey (XY to ZW or vice versa) are more likely to be favoured by selection if there is selection upon both haploid and diploid genotypes rather than diploid selection alone. 
This prediction could be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
In animals, one might expect gametic competition to be stronger in species where sperm is required to live for a long time after spermatogenesis because transcripts shared during spermatogenesis may become depleted, revealing the haploid phenotype of the sperm \citep{Immler:2014im}. 
Given the caveats mentioned above about the form of meiotic drive modelled, we would also expect that heterogametic transitions in sex determination would be more common in clades where there is meiotic drive.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Haplotype invasion region plot
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] () {
	\includegraphics[width=0.5\linewidth]{regionplot_2a}
	\includegraphics[width=0.5\linewidth]{regionplot_hap}
	};
\end{tikzpicture}
\caption{
Gray region depicts the relative fitnesses of male homozygotes, $w_AA^\male$ and $w_aa^\male$, that allow both neo-W haplotypes to invade, from any stable resident equilibria (equations \ref{eq:tightequil}).
In the white region neo-W haplotypes paired with the allele fixed on the Y cannot invade.
Parameters as in \cite{Otto2014} Figure 2a: $w_A^\female=w_a^\female=w_A^\male=w_a^\male=1$, $\alpha^\female=\alpha^\male=1/2$, $w_Aa^\female=w_Aa^\male=1$, and $w_AA^\female=w_aa^\female=0.75$, with $r=0$.
}
\label{fig:regionplot_ 2a}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%\textcolor{blue}{
In Figure \ref{fig:regionplot_ 2a}A we show the region of parameter space within which both neo-W haplotypes invade ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) when there is overdominance in females and no haploid selection \citep[corresponding to Figure 2a in][]{Otto2014}. 
Wherever both haplotypes have positive growth rates (gray region of Figure \ref{fig:regionplot_ 2a}), invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where only one haplotype can spread (white region of Figure \ref{fig:regionplot_ 2a}), a neo-W can invade as long as equation \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is small enough and yet still indicates that more loosely linked sex-determining regions can spread.
It is also possible for haploid selection to drive the invasion of a loosely linked neo-W. 
Take for instance selection directionally favouring $A$ in both diploid sexes and meiotic drive in males.
Figure \ref{fig:regionplot_ 2a}B then shows that ploidally-antagonisitic selection can allow both neo-W haplotypes to invade. 
%}

@@

%\textcolor{blue}{As with the discussion, this section might need some editing for (1) unclearly describing looser linkage never evolving without haploid selection, which we now show is possible above and (2) going on about sex ratio selection too much - this should be changed to something along the lines of  ``male biased zygotic sex ratios ($\zeta-1/2$) are of the order of $\epsilon^2$. Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
%Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
%Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section.
%''
%}

@@

%Sex ratios
%Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
%Our model provides a good opportunity to determine whether Fisherian sex-ratio selection provides a useful explanation for the evolution of new sex-determining loci in other contexts. 
%Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
%We will also disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
%In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
%Thus, if the zygotic sex ratio is crucial to the evolution of new genetic sex-determining systems, invasion into ZW and XY systems will be distinct. 
%However, under weak selection we find that invasion by a homogametic neo-sex-determining allele (XY to XY or ZW to ZW) or by a heterogametic neo-sex-determining allele (XY to ZW or ZW to XY) occur under the same conditions. 
%That is, we can show that $\lambda_{Y',XY}=\lambda_{W',ZW}$ and $\lambda_{Y',ZW}=\lambda_{W',XY}$ (at least up to order $\epsilon^3$; for a numerical example, compare Figure \ref{fig:Combination_Turnover}A,B to Figure \ref{fig:Combination_Turnover}C,D).
%As it turns out, under weak selection the strength of sex-ratio selection favouring, say, the invasion of a neo-W in an XY system is the same as the strength of meiotic drive favouring the invasion of a neo-Y in a ZW system. 
%Even when these forces are not exactly the same (e.g., under tight sex-linkage; compare black and red curves near -25 and 25cM in Figure \ref{fig:Combination_Centimorgans}), it is important to remember that sex-ratio selection is only one of many potential selective forces acting to determine transitions between sex-determining systems.
%It is even possible for the other selective forces to overwhelm sex-ratio selection and favour sex-determination transitions that create sex-ratio biases (Figure \ref{fig:Combination_Turnover}A,C).

Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
Disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
We might therefore expect a difference in the potential for XY to ZW and ZW to XY transitions. 
However, with weak selection we find that invasion by a neo-W into an XY system and invasion by a neo-Y into a ZW system occur under the same conditions. 
As it turns out, with weak selection, the strength of a male biased zygotic sex ratio favouring the invasion of a neo-W is equal to the strength of meiotic drive favouring the invasion of a driving neo-Y are equal (at least up to order $\epsilon^2$).
As selection becomes stronger (or linkage becomes tighter), this symmetry between sex-ratio selection and drive is lost, causing differences in the strength of selection favouring the two heterogametic transitions (compare red to black near -25cM and 25 cM in Figure \ref{fig:Combination_Centimorgans}).

%Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
%Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
%Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section

@@

We next consider the case where the new sex-determining mutation, $m$, causes sex to be determined probabilistically or by heterogeneous environmental conditions (environmental sex determiner, ESD). 
We assume that individuals carrying the $m$ allele develop as females with probability $k$ (e.g., in a fraction $k$ of the environments they randomly experience).
The characteristic polynomial determining the eigenvalues of the 8 equation system (equations \ref{eq:recursions}) does not reduce for ESD mutants as it does for $k=0$ or $k=1$. 
We therefore focus on weak selection here. 
Assuming weak selection, the spread of these mutations is given by

@@

%\textcolor{blue}{maybe re-order to put the results up front and focus on sex ratio less. Somewhere around these first two paragraphs we should include our tight linkage results. At the moment it doesn't make total sense because it assumes that tighter linkage is always favoured unless there is haploid selection.}

%OPENING
Two predominant theories explaining the remarkably high frequency of transitions between sex-determination systems are sexually-antagonistic selection and sex-ratio selection \citep[reviewed in][]{Blaser2012}.
The former predicts that neo-sex-determining alleles can invade when they arise in closer linkage with a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
The latter predicts that neo-W alleles will invade an XY system when there is a male bias caused by haploid selection in males, and vice-versa, a neo-Y will invade a ZW system when there is a female bias caused by haploid selection in females \citep{Kozielska:2010vm,Ubeda:2015fx}.
Here we have shown that both predictions must be amended when recombination is weak relative to selection or selection happens in both diploid and haploid phases.

%TIGHT LINKAGE RESULTS
When the rate of recombination between the ancestral sex-determining locus and a locus under selection is small relative to the strength of selection (i.e., sex-linkage is tight), heterogametic transitions (XY to ZW or ZW to XY) that reduce sex-linkage are possible, with or without haploid selection (Figure \ref{fig:regionplot_2a}A) or sexually-antagonistic selection (Figure \ref{fig:regionplot_2a}B).
The likelihood of these transitions are driven by sex-ratio selection, direct selection on alleles linked to the neo-sex-determining allele, the ability of the neo-sex-determining allele to avoid selection in one sex, and the ability of the neo-sex-determining allele to bring alleles on the sex-specific chromosome in the ancestor into the other sex (given that the neo-sex determining allele is epistatically dominant to its predecessor).
This possibility was overlooked in \cite{vanDoorn:2010hu}, possibly because they did not explicitly calculate the resident equilibria \citep[equation \ref{eq:tightequil};][]{Lloyd1977,Otto2014}.
Interestingly, there is substantial overlap between the parameter space that allows both neo-W-$A$ and neo-W-$a$ haplotypes to spread in an XY system and that which selects for increased recombination between X and Y chromosomes \citep[e.g., compare gray region in Figure \ref{fig:regionplot_2a}A with coloured regions of Figure 2(a) in][]{Otto2014}.
This makes sense as when both neo-W haplotypes can spread the neo-W can invade despite reducing sex-linkage, i.e., the rate of recombination between the sex-determining allele and the selected locus increases.
\textcolor{red}{explain difference too? also explain why a neo-Y doesn't invade to reduce linkage despite selection for more recombination? mention something about haploid selection (Fig7B)?}

%HAPLOID SELECTION CAN ACT LIKE SEXUALLY-ANTAGONISTIC SELECTION
%It has previously been demonstrated that new sex-determining systems can evolve if there is genetic variation maintained by sexually-antagonistic selection \citep{vanDoorn:2007eu,vanDoorn:2010hu}. 
%In particular, 
Under weak selection (i.e., loose sex-linkage), transitions to new sex-determining systems can occur when new sex-determining regions are more closely linked to a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
Our results show that genetic variation at loci that experience haploid selection can also generate selection in favour of new sex-determining systems. 
New sex-determining alleles are again favoured if they are more closely linked to a locus under haploid selection.
%HAPLOID SELECTION ALSO ALLOWS LOOSER LINKAGE TO EVOLVE
However, with haploid selection, heterogametic transitions (XY to ZW or ZW to XY) can also occur when the new sex-determining region is less closely linked to the locus under selection. 
Neo-W (neo-Y) alleles invade when their fitness in females (males) is greater than the mean fitness of females (males) under the ancestral sex-determination system.
With sexually-antagonistic selection (between diploid sexes) only, linkage between a selected locus and the sex-determining region strengthens associations between male beneficial alleles and the male-determining allele (Y or Z) and between female beneficial alleles and the female-determining allele (X or W). 
Thus, the mean fitness of both males and females increases with closer linkage to the sex-determining region. 
Therefore, new sex-determining alleles only invade if they are more closely linked than the ancestral sex-determining region. 
However, if there is haploid selection on loci linked to an XY (ZW) sex-determining region, selection can maintain polymorphisms at which the mean fitness of females (males) is lower than it would be without sex-linkage. 
In these cases, unlinked neo-W (neo-Y) alleles can increase female (male) fitness, at a cost to the other sex, and invade despite lowering mean fitness (Figure \ref{fig:Combination_MeanFit}).

%%SEX-RATIO SELECTION
%\textcolor{blue}{maybe drop paragraph?}
%\textcolor{red}{I messed with the sex-ratio selection paragraphs to tone down our "it doesn't matter" speech from before. Have at any amendments you'd like to make.}
%Linkage between haploid selected loci and sex-determining regions cases biased zygotic sex ratios \citep{Hamilton:1967ts,Burt:2006,Field:2012fd,Field:2013cc}.
%One might then expect Fisherian sex-ratio selection to drive the spread of new sex-determining systems that bring the sex ratio closer to 50:50. 
%Fisherian sex-ratio selection follows from the fact that, for an autosomal locus, half of the genetic material is inherited from a male, and half from a female \citep{Fisher:1930wy,West:2009we}. 
%Thus, if the population sex ratio is biased towards females, the average per-individual contribution of genetic material to the next generation from males is greater than the contribution from females (and vice versa for male-biased sex ratios). 
%Therefore, a mutant that increases investment in males will spread via the higher per-individual contributions made by males. 
%%%That is, under Fisherian sex-ratio selection, the success of a mutant relative to the non-mutant depends, in equal parts, on the contributions made by males and females to the next generation. 
%An implicit assumption of Fisherian sex-ratio selection is that the mutant allele is autosomal and has the same inheritance pattern as the non-mutant allele. 
%The mutations we consider here, neo-sex-determining alleles, break this assumption. 
%For example, the success of neo-Y/neo-W mutations depends only on the number of alleles contributed by males/females (Table \ref{tab:haplotype_growth}). 
%In this respect, a neo-W is similar to a cytoplasmic element, which also does not experience selection to balance sex ratios \citep{Frank:1989vl,Werren:1998co,Chase:2007bf}.
%Even mutants that are equally likely to be found in males or females, such as an environmental sex determination mutation (equation \ref{eq:lambda_ESD}), are not strictly autosomal if they determine sex. 
%Thus, despite the fact that sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal modifiers \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va}, we do not find evidence of Fisherian sex-ratio selection acting during invasion by neo-sex-determination systems (e.g., see Figure \ref{fig:Combination_Turnover} and \citealt{Ubeda:2015fx}, in which a neo-Y invades despite biasing sex ratios).

%\textcolor{blue}{Maybe we could go along the angle of ``under the weak recombination assumptions, there is no asymmetry between XY to ZW and ZW to XY, indicating that sex ratio selection does not dominate. Asymmetry can be larger when linkage is tight (e.g., Fig 2C) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
%Thus, haploid selection can favour heterogametic transitions both via sex ratio selection and via the fitness of alleles that are associated with the neo-sex-determining allele; these selection pressures are often of equal magnitude.''}

Sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va} and sex-linked \citep{Ubeda:2015fx} modifiers.
We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
For instance, when an allele that drives in males is linked to an XY locus it will often become associated with the Y and therefore produce a male bias ($\zeta<1/2$).
This male bias increases the potential for a neo-W to invade (as we then have $(2\zeta)^{-1}>1$, Table \ref{tab:haplotype_growth}), which can equalize the sex-ratio \citep[for a related example see][]{Ubeda:2015fx}.
%However, the neo-W may not equalize the sex-ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
However, this sex-ratio selection can be overwhelmed when the driving allele has additional selective effects (e.g., when it is detrimental to male diploids but beneficial for female diploids; Table \ref{tab:specialcases}), preventing the neo-W from invading. %and equalizing the sex ratio.
Conversely, these additional selective effects can even favour transitions between sex-determining systems that create new sex-ratio biases.
For example, in an ancestral ZW system, an allele that drives only in males can allow a linked neo-Y to invade, despite the fact it creates a male bias.
This of course generates new sex-ratio selection that may drive further turnover \citep{Ubeda:2015fx}.
%What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.
Furthermore, with weak selection/loose sex-linkage, there is no asymmetry between XY to ZW and ZW to XY transitions, indicating that sex-ratio selection does not dominate (i.e., the sex-ratio bias created by haploid selection impacts the spread of a neo-W into an XY system the same way it impacts the spread of a neo-Y into an unbiased ZW system). 
An asymmetry can develop when sex-linkage is tight (e.g., Figure \ref{fig:Combination_Centimorgans} near -25cM and 25cM) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
Thus, haploid selection can favour heterogametic transitions both via sex-ratio selection and via the fitness of alleles that are associated with the neo-sex-determining allele; and these selection pressures are often of equal magnitude.

@@

%HOW TO TEST EMPIRICALLY
\textcolor{blue}{I think Sally was sceptical of this idea. i.e., if the second term in \ref{eq:lambda_neoW} is negative, then get less transitions with haploid selection included, right? }
\textcolor{red}{Yeah, and given that only 1/2 the discovered plant sex chromosomes are hetermorphic and there are so few plant sex chromosomes to begin with (Charlesworth 2016 Annual Review Plant Biology), might it be the case that haploid selection is constraining things? perhaps we could think of some particular forms of haploid selection that might be common and give their predicted effects, or just say some things that are possible with haploid selection that are not without it -- if there are any. just looking at eq 3 though, its gonna be hard to get a low freq of A on the Y if haploid selection in males favours A, i.e., it should often be positive when there is no female haploid selection, right?}
Taken at face value, our results indicate that transitions in heterogametey (XY to ZW or vice versa) are more likely to be favoured by selection if there is selection upon both haploid and diploid genotypes rather than diploid selection alone. 
This prediction could be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
In animals, one might expect gametic competition to be stronger in species where sperm is required to live for a long time after spermatogenesis because transcripts shared during spermatogenesis may become depleted, revealing the haploid phenotype of the sperm \citep{Immler:2014im}. 
Given the caveats mentioned above about the form of meiotic drive modelled, we would also expect that heterogametic transitions in sex determination would be more common in clades where there is meiotic drive.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Haplotype invasion region plot
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] () {
	\includegraphics[width=0.5\linewidth]{regionplot_2a}
	\includegraphics[width=0.5\linewidth]{regionplot_hap}
	};
\end{tikzpicture}
\caption{
Parameter space (gray) where both neo-W haplotypes can invade from the same stable resident $r=0$ equilibria (equations \ref{eq:tightequil}), and therefore where an unlinked neo-W can invade an XY system with perfect sex-linkage.
\textbf{A}, In the absence of haploid selection, both neo-W haplotypes can invade for much of the parameter space where the relative fitnesses of male homozygotes, $w_{AA}^\male$ and $w_{aa}^\male$, are both less than that of the heterozygote, $w_{Aa}^\male = 1$.
In the white region neo-W haplotypes paired with the allele fixed on the Y cannot invade.
Parameters as in \cite{Otto2014} Figure 2a: $w_A^\Hermaphrodite=w_a^\Hermaphrodite$, $\alpha^\Hermaphrodite=1/2$, $w_Aa^\Hermaphrodite=1$, and $w_AA^\female=w_aa^\female=0.75$.
\textbf{B}, When selection is the same in both diploid sexes ($w_{aa}^\Hermaphrodite = 1$, $w_{Aa}^\Hermaphrodite = 1 + h s$, $w_{AA}^\Hermaphrodite = 1+s$), both neo-W haplotypes can invade over a portion of the parameter space where selection in diploids ($s$) opposes the force of drive during meiosis in males ($\alpha_{\Delta}^\male$).
Parameters: $w_A^\Hermaphrodite=w_a^\Hermaphrodite$, $\alpha^\female=1/2$, $h=1/2$.
}
\label{fig:regionplot_2a}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
